Models

### 2021-01-25 analysis/05_model_metrics.Rmd
Get p-values for models, plot dornelas figures

Fit GLM to overall metric trend
```{r}
# model for overall trend (random effect w/varying intercept)
## should be GLS according to Dornelas supplement, but glmer isn't happy
metric_mean_fit <- metrics %>%
  filter(!is.na(value)) %>%
  group_nest(metric) %>%
  mutate(model = map(data, ~lmer(logvalue ~ year + (1|study_id) + (1|rarefyID), data = .x)))

@@

@@

```{r}
indiv_ts_fit <- metrics %>% 
  filter(!is.na(value)) %>%
  group_by_at(setdiff(names(metrics), c("year", "value"))) %>%
  filter(n_distinct(year) > 1) %>% #filter out rarefyID's that don't have multiple years after dropping Na's
  group_nest() %>%
  mutate(model = map(data, ~ lm(value ~ year, data = .x)))

@@

Plot of S and diversity metrics
```{r}
indiv_ts_coef %>%
  filter(term == "year", metric %in% c("FRic", "FEve", "FDiv", "FDis", "S")) %>%
  select(metric, estimate) %>% 
  ggplot() +
  geom_histogram(aes(x = estimate)) + #,
                 #binwidth = 2) +
  scale_x_continuous(name = 'slope') +
  #scale_y_continuous(breaks = c(0,2500,5000,7500,10000,12500, 15000, 17500, 20000,22500)) +
  labs(tag = 'B',
       y = 'Number of cells') +
  theme_bw() +
  theme(panel.grid.minor.x = element_blank()) +
  facet_wrap(vars(metric), scales = "free")
```

>>>>

Fit GLM to overall metric trend
```{r}
# model for overall trend (random effect w/varying intercept)
## should be GLS according to Dornelas supplement, but glmer isn't happy
metric_mean_fit <- metrics %>%
  filter(!is.na(value)) %>%
  mutate(minyear = min(year), maxyear = max(year)) %>%
  group_nest(metric, minyear, maxyear) %>%
  #mutate(model = map(data, ~lmer(logvalue ~ year + (1|study_id) + (1|rarefyID), data = .x)),
  mutate(model = map(data, ~lmer(logvalue ~ year + (1|rarefyID), data = .x)),
         prediction = map(model, ~predict(.x), newdata = as.data.frame(year = minyear:maxyear)))

@@

mean_tests <- metric_mean_fit %>%
  #filter(metric %in% c("FRic", "FEve", "FDiv", "FDis", "S")) %>%
   mutate(p_value = map(model, ~drop1(.x, test = "Chisq") %>% select(`Pr(Chi)`) %>% slice(2) %>% pull())) %>% 
  select(metric, p_value) %>%
  left_join(metric_mean_coef %>% 
              filter(term == "year") %>% 
              select(metric, estimate))

@@

```{r}
indiv_ts_fit <- metrics %>% 
  filter(!is.na(value)) %>%
  group_by_at(setdiff(names(metrics), c("year", "value", "logvalue"))) %>%
  filter(n_distinct(year) > 1) %>% #filter out rarefyID's that don't have multiple years after dropping Na's
  group_nest() %>%
  mutate(model = map(data, ~ lm(logvalue ~ year, data = .x)))

@@

Plot of S and diversity metrics
```{r}
indiv_ts_coef %>%
  filter(term == "year", metric %in% c("FRic", "FEve", "FDiv", "FDis", "S")) %>%
  select(metric, estimate) %>% 
  ggplot() +
  geom_histogram(aes(x = estimate)) + #,
                 #binwidth = 2) +
  scale_x_continuous(name = 'slope') +
  #scale_y_continuous(breaks = c(0,2500,5000,7500,10000,12500, 15000, 17500, 20000,22500)) +
  labs(tag = 'B',
       y = 'Number of cells') +
  theme_bw() +
  theme(panel.grid.minor.x = element_blank()) +
  facet_wrap(vars(metric), scales = "free")
```
Reproduce Dornelas figure - all data points & regressions w/overall mean
```{r}
#get climate classification
metadata <- read_csv(system.file("extdata", "biotime/biotime_metadata.csv", package = "biodivTS"))

fric_trend <- metrics %>%
  filter(metric == "FRic") %>%
  ggplot(aes(x = year, y = logvalue)) +
  # geom_smooth(aes(color = climate, group = plot_id), method = "glm", se = FALSE) +
  # geom_point(aes(color = climate, group = plot_id)) +
  geom_smooth(aes(group = rarefyID), color = "grey", method = "glm", se = FALSE) +
  geom_point(aes(group = rarefyID) ,color = "grey") +
  geom_smooth(color = "black", method = "glm", se = FALSE) +
  theme(legend.position = "none") 
ggsave("fric_trends.png", fric_plot)
```

```{r}
library(lme4)

plot_metric <- function(metric_name, ylabel){

  data <- metrics %>%
    filter(metric == metric_name, !is.na(value), !is.infinite(value))

  model_coef <- metric_mean_coef %>%
    filter(metric == metric_name) %>%
    select(term, estimate)

  intercept <- model_coef %>% filter(term == "(Intercept)") %>% pull(estimate)
  slope <- model_coef %>% filter(term == "year") %>% pull(estimate)

  pred <- data.frame(year = min(data$year):max(data$year)) %>%
    mutate(pred = intercept + year * slope)

  trend_plot <- data %>%
    left_join(pred, by = "year") %>%
    #left_join(metadata %>% select(study_id, climate)) %>%
    ggplot(aes(x = year, y = logvalue)) +
    # geom_smooth(aes(color = climate, group = rarefyID), method = "glm", se = FALSE) +
    # geom_point(aes(color = climate, group = rarefyID)) +
    geom_point(aes(group = rarefyID) ,color = "grey", size = 0.30) +
    geom_smooth(aes(color = climate_mod, group = rarefyID), method = "glm", se = FALSE, size = 0.5) +
    #geom_smooth(color = "black", method = "glm", se = FALSE) +
    geom_line(mapping = aes(y = pred)) +
    theme(legend.position = "none") +
    theme_classic() +
    ylab(ylabel) +
    labs(color = "Climate")

  ggsave(paste0(metric_name, "_trend.png"), trend_plot)
  return(trend_plot)
}

plot_metric("FRic", "log(FD)")

plot_map <- tibble(metric_name = c("S", "FRic", "FEve", "FDiv", "FDis"), 
           ylabel = c("log(S)", "log(FD)", "log(Evenness)", "log(Divergence)", "log(Dispersion)"))
pmap(plot_map, plot_metric)
```

```{r}

```
